On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-26 TTau_Halpha/kurosawa.tex
end of thursday - th
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@637 25464f21-6099-402c-b30a-640f1d40d38b

    \includegraphics[%
      clip,
      scale=0.45]{figures/discwind_hybrid_ha_image.eps}

@@

\subsection{Classification scheme proposed by Reipurth et. al (1996)}

@@

  \caption{Sample H$\alpha$ model profiles which characterise the
    classification by \citet{reipurth:1996}. The combination of
    magnetospheric accretion flow, the accretion disc, and the disc wind
    model can reasonably reproduce wide ranges of H$\alpha$ profiles seen
    in observations.  The model parameters used for each profile and
    short comments are summarised in
    Table~\ref{tab:classification}. The horizontal axes 
    are velocities in $\kmps$, and the vertical axes are normalised
    flux.}

@@

\citet{reipurth:1996} proposed the two-dimensional classification
of H$\alpha$ emission profiles of T~Tauri stars and Herbig Ae/Be
stars. Their classification scheme contains four classes (I, II, III
and IV) differentiated by the ratio of the secondary-to-primary emission
components in the profiles. Each classes are divided into two sub-classes
(B and R) which depends whether the absorption component is on the
blue or red side. Readers are refereed to Fig.~4 of their paper.
Figure~\ref{fig:atlas_reipurth} shows the sample model profiles
which are classified according to the definition of \citet{reipurth:1996}.
The combination of the disc wind, magnetospheric accretion, and
accretion disc can reasonably reproduce all the classes of the profiles
seen in observations. The corresponding parameters of the
disc-wind-magnetosphere hybrid model used to reproduce the profiles
in the figure are summarised in Table~\ref{tab:classification} along
with brief comments on possible physical conditions which leads to
the profiles in each class. In the following, we will discuss each 
classification type of \citet{reipurth:1996} based on our model
results.

\textbf{Type~I}: 
Using the samples of H$\alpha$ from 43 CTTS, \citet{reipurth:1996}
found this type to be the second most common (26~per~cent) among
them. Type~I (symmetric around line centres), is mainly
produced in the magnetosphere.  Although not shown here, this type of
profile can be also reproduced with the disc-wind-magnetosphere hybrid
model with low inclination angles ($\lesssim 10^\circ$).  Since there is
no material in the polar region in the disc wind model, the observer
can look the magnetosphere almost directory for high inclination
cases, provided that wind temperature is low enough so that the wind
emission is negligible.

\textbf{Type~II-B}:  
This type can be produced by the disc-wind-magnetosphere hybrid models
at moderate to high inclination angles with wide ranges of the wind
acceleration rate ($0.2 \lesssim \beta  \lesssim 2.0$) and the wind
temperature ($7000\,\mathrm{K} \lesssim T \lesssim 9000 \mathrm{K}$). If
the inclination is high ($i \sim 80^\circ$), the disc-wind
acceleration parameter should be relatively small ($\beta \lesssim
0.5$). If the inclination is moderate ($i \sim 60^\circ$), the wind
acceleration parameter should be relatively high ($\beta \gtrsim 2.0$).

\textbf{Type~II-R}:  
The physical conditions which result in this type
is very similar to those of Type~II-B according to the
disc-wind-magnetosphere hybrid models. The main difference is between  
the two is in their inclination angles. This type is more likely
seen at higher inclination angles ($i \sim 80^\circ$) compared to Type
II-B.  Commonly seen in the models with the disc wind only model.

\textbf{Type~III-B}:
This type is the most commonly seen profiles in the CTTS H$\alpha$
samples in \citet{reipurth:1996}. Interestingly, the most common
profile type seen in the disc-wind-magnetosphere hybrid model are
Type~III-B profiles (c.f.~Figures~\ref{fig:atlas_hybrid} and
\ref{fig:atlas_hybrid}). In this type of profiles, the
main emission comes from the magnetosphere, and the profile shape is
altered from a slightly blue-asymmetric one to a slightly
red-asymmetric one via the wind absorption in the blue wing.  In the
models with the disc-wind, the acceleration parameter should be
relatively small ($\beta<0.5$). The profile type changes from Type
III-R to Type~II-R if the value of $\beta$ increases. The
inclination angles must be moderate to high ($ 60^\circ \lesssim i \lesssim
 80^\circ$).

\textbf{Type~III-R}: 
This is the least common type of the observed profile (2~per~cent)
according to \citet{reipurth:1996}. Incidentally, we also found it
difficult to reproduce this type of profile since this can be achieve
only at narrow range of parameter space (very high inclination angle
with a disc wind model, $T_{\mathrm{max}}\sim8300\,\mathrm{K}$, and
$\dot{M}_{\mathrm{acc}}\sim10^{-8}\,\mathrm{M_{\sun}yr^{-1}}$).
Since the inclination angles must be high, this
indicates that the accretion disc plays an important role in producing
this type of the profile, but not the wind.  The rareness of this type in
the observation can be understood by the required high inclination
angle (almost edge-on) in our models.

\textbf{Type~IV-B}: 
This type of the profiles is very similar to Type III-B, but the wind
in this type of profile may be accelerating faster (a smaller $\beta$
value) than that in Type~III-B.  In the models with the disc-wind,
this type is seen at moderate to high inclinations ($60^\circ \lesssim
 i \lesssim  80^\circ$) with relatively small value of the acceleration
parameters ($\beta \lesssim 0.5$).

\textbf{Type~IV-R}: 
This is the second least common type of the observed profiles (5~per~cent)
which shows the inverse P-Cyg profile shape.  This can be understood
since the magnetospheric accretion model of \citet{hartmann:1994}
predicts that the redshifted absorption can occur at a limited range
of inclination angle ($i\simeq 50^\circ$, but depends on the geometry of
the magnetosphere) with which the hot continuum flux from the
footprint of the magnetosphere is seen through relatively fast moving
gas in the accretion stream.  The wind component must be very
weak. The mass-accretion rate should be much lower
($\dot{M}_{\mathrm{acc}} \sim 10^{-9}\,\mathrm{M_{\sun}\,yr^{-1}}$)
than that of the models which produce Type~I profiles since the flux
in the broadening wings can not be too high to produce the inverse
P-Cyg profile.

>>>>

    \includegraphics[%
      clip,
      scale=0.45,angle=270]{figures/discwind_hybrid_ha_image.eps}

@@

\subsection{The morphological classification
 scheme proposed by Reipurth et. al (1996)}

@@

  \caption{Sample H$\alpha$ model profiles which characterise the
    classification scheme of \citet{reipurth:1996}. The combination of
    magnetospheric accretion flow, the accretion disc, and the disc-wind
    can reproduce the wide range of H$\alpha$ profiles seen
    in observations.  The model parameters used for each profile along with
    short comments are given in
    Table~\ref{tab:classification}. The horizontal axes 
    are velocities in $\kmps$, and the vertical axes are normalised
    flux.}

@@

\citet{reipurth:1996} proposed a two-dimensional classification
of H$\alpha$ emission profiles of T~Tauri stars and Herbig Ae/Be
stars. Their classification scheme contains four classes (I, II, III
and IV) differentiated by the ratio of the secondary-to-primary emission
components in the profiles. Each class is divided into two sub-classes
(B and R) which depends whether the secondary peak is on the
blue or red side of the primary peak. Readers are refereed to Fig.~4 of their paper.
Figure~\ref{fig:atlas_reipurth} shows  sample model profiles
which are classified according to the definition of \citet{reipurth:1996}.
The combination of the disc wind, magnetospheric accretion flow, and
the accretion disc can reproduce all the classes of the profiles
seen in observations. The corresponding parameters of the
disc-wind-magnetosphere hybrid model used to reproduce the profiles
in the figure are summarised in Table~\ref{tab:classification} along
with brief comments on possible physical conditions which lead to
the profiles in each class. Below we discuss the Reipurth scheme in the
context of our models parameters.

\textbf{Type~I}:

\citet{reipurth:1996} found that Type I profiles (symmetric around line centres) constituted 26~per~cent
of the 43 CTTS H$\alpha$ profiles in their sample, making it the second most common
morphological type. We find that Type~I  profiles are dominated by magnetospheric
emission at rates in excess of $\appox 10^{-8}\, \dot{M_\odot}\,{\rm yr}^{-1}$. Type~I profiles
appear at a wide range of inclinations (Figure~\ref{magnetosphere only inc dependence figure}).
An object that displays a Type~I profile may have relatively weak disc wind, but in this
case it must be viewed from the pole (in order that no wind material is seen projected
against the stellar photosphere).

\textbf{Type~II-B}:

These profiles have a secondary blue peak that is in excess of half the strength of the 
primary peak, and comprised 16~per~cent of the Reipurth sample.

This type can be produced by the disc-wind-magnetosphere hybrid models
at moderate to high inclination angles with wide ranges of the wind
acceleration rate ($0.2 \lesssim \beta  \lesssim 2.0$) and the wind
temperature ($7000\,\mathrm{K} \lesssim T \lesssim 9000 \mathrm{K}$). If
the inclination is high ($i \sim 80^\circ$), the disc-wind
acceleration parameter should be relatively small ($\beta \lesssim
0.5$). If the inclination is moderate ($i \sim 60^\circ$), the wind
acceleration parameter should be relatively high ($\beta \gtrsim 2.0$).

\textbf{Type~II-R}:

These

The physical conditions which result in this type
is very similar to those of Type~II-B according to the
disc-wind-magnetosphere hybrid models. The main difference is between  
the two is in their inclination angles. This type is more likely
seen at higher inclination angles ($i \sim 80^\circ$) compared to Type
II-B.  Commonly seen in the models with the disc wind only model.

\textbf{Type~III-B}:

xxx describe this type xxxx

This type is the most commonly seen profiles in the CTTS H$\alpha$
samples in \citet{reipurth:1996}. Interestingly, the most common
profile type seen in the disc-wind-magnetosphere hybrid model are
Type~III-B profiles (c.f.~Figures~\ref{fig:atlas_hybrid} and
\ref{fig:atlas_hybrid}). In this type of profiles, the
main emission comes from the magnetosphere, and the profile shape is
altered from a slightly blue-asymmetric one to a slightly
red-asymmetric one via the wind absorption in the blue wing.  In the
models with the disc-wind, the acceleration parameter should be
relatively small ($\beta<0.5$). The profile type changes from Type
III-R to Type~II-R if the value of $\beta$ increases. The
inclination angles must be moderate to high ($ 60^\circ \lesssim i \lesssim
 80^\circ$).

\textbf{Type~III-R}:

xxx describe this type xxxx

This is the least common type of the observed profile (2~per~cent)
according to \citet{reipurth:1996}. Incidentally, we also found it
difficult to reproduce this type of profile since this can be achieve
only at narrow range of parameter space (very high inclination angle
with a disc wind model, $T_{\mathrm{max}}\sim8300\,\mathrm{K}$, and
$\dot{M}_{\mathrm{acc}}\sim10^{-8}\,\mathrm{M_{\sun}yr^{-1}}$).
Since the inclination angles must be high, this
indicates that the accretion disc plays an important role in producing
this type of the profile, but not the wind.  The rareness of this type in
the observation can be understood by the required high inclination
angle (almost edge-on) in our models.

\textbf{Type~IV-B}:

xxx describe this type xxxx

This type of the profiles is very similar to Type III-B, but the wind
in this type of profile may be accelerating faster (a smaller $\beta$
value) than that in Type~III-B.  In the models with the disc-wind,
this type is seen at moderate to high inclinations ($60^\circ \lesssim
 i \lesssim  80^\circ$) with relatively small value of the acceleration
parameters ($\beta \lesssim 0.5$).

\textbf{Type~IV-R}:

xxx describe this type xxxx

This is the second least common type of the observed profiles (5~per~cent)
which shows the inverse P-Cyg profile shape.  This can be understood
since the magnetospheric accretion model of \citet{hartmann:1994}
predicts that the redshifted absorption can occur at a limited range
of inclination angle ($i\simeq 50^\circ$, but depends on the geometry of
the magnetosphere) with which the hot continuum flux from the
footprint of the magnetosphere is seen through relatively fast moving
gas in the accretion stream.  The wind component must be very
weak. The mass-accretion rate should be much lower
($\dot{M}_{\mathrm{acc}} \sim 10^{-9}\,\mathrm{M_{\sun}\,yr^{-1}}$)
than that of the models which produce Type~I profiles since the flux
in the broadening wings can not be too high to produce the inverse
P-Cyg profile.
